\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgments}

I would like to thank Prof.\ Luc Van Gool for giving me the opportunity to
pursue a PhD in his lab. Thanks to him, I was able to freely choose research
directions that interested me, while being financially supported.  
I would like to thank Radu Timofte for believing in me and supporting my work.
The environment of the lab helped me to stay focused and motivated, and I spent
many great hours there.

This thesis would not have been possible without my two close collaborators,
Eirikur Agustsson and Michael Tschannen.  I met Michael when I was doing
my master, where he supervised my second semester thesis. Supervised by him and
Eirikur, I wrote my master thesis, and since then our collaboration has been
steadily going throughout the PhD. Without their insightful comments and
relentless persistence, this work would not have been possible. 

I would like to thank Sergi Caelles, for making the days at the lab brighter
with discussions, and for helping me with tough decisions. I would like to
thank Kevis Maninis for helping me settle into PhD life with his advice early
in the studies.
The final work presented in this thesis was done during an internship at
Google, and I would like to thank George Toderici for the many hours he spent
supporting the paper. Without him, we would not have the thorough user study
that we have now, and the paper would be overall less solid.
I would like to thank Luisa Blom for supporting me mentally
through many tough deadlines.
I would like to thank my family for the thorough support throughout
my studies. Thanks to them, I was able to focus on my studies.



\endgroup
